---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# shayhai

<!-- badges: start -->
<!-- badges: end -->

## Introduction

shayhai is a teaching-focused R package for exploring the burden of healthcare-associated infections (HAIs) using simulated datasets. It provides tidy data, helper functions, and an interactive Shiny app to summarise and visualise HAI burden — including infection counts, attributable deaths, and disability-adjusted life years (DALYs).

This package is designed for coursework, demos, and data communication exercises.

## Usage

You can install the development version of shayhai from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ETC5523-2025/assignment-4-packages-and-shiny-apps-Shayna-Yang")
```
Load the package:

  ```{r}
  library(shayhai)
  ```

### Data Included

The package ships with two small simulated datasets:

- `shayhai_cases` — German PPS sample (2011)

- `shayhai_cases_ecdc` — EU/EEA PPS sample (2011–2012)

Each row represents a simulated “case” weighted to reflect the national burden.
Key variables include:

- infection_type: infection category (`HAP`, `SSI`, `BSI`, `UTI`, `CDI`)

- `weight_pop`: population weight for national scaling

- Demographic and infection descriptors (used in filtering and plotting)

Example preview:

```{r}
head(shayhai_cases)
```

### Shiny App

The package includes an interactive dashboard to explore infection burden by type, survey module, and demographic group.

App features:

- Bubble plot: HAIs vs deaths, bubble size = DALYs

- Bar plot: Compare German PPS vs ECDC PPS (with 95% uncertainty intervals)

- Age pyramid: DALYs by sex and age group

You can explore the interactive [Shiny app](https://etc5523-2025.github.io/assignment-4-packages-and-shiny-apps-Shayna-Yang/articles/Examples.html) included in the package.

## Data Sources

The simulated datasets mirror the structure of two major PPS surveys:

- German PPS (2011): 46 hospitals, 9,626 patients

- EU/EEA PPS (2011–2012): 273,753 patients across 29 countries

Metrics (HAIs, deaths, DALYs, YLL, YLD) follow the methodology of the Burden of Healthcare-Associated Infections (BHAI) framework as described in:

Zacher B, Haller S, Willrich N, et al. Application of a new methodology and R package reveals a high burden of healthcare-associated infections (HAI) in Germany compared to the average in the European Union/European Economic Area, 2011 to 2012. Euro Surveill. 2019;24(46):1900135. [https://doi.org/10.2807/1560-7917.ES.2019.24.46.1900135](https://doi.org/10.2807/1560-7917.ES.2019.24.46.1900135)

These data are simplified for educational use in ETC5523: Communicating with Data.

## Documentation

Full documentation and examples:

**Site home: **[https://ETC5523-2025.github.io/assignment-4-packages-and-shiny-apps-Shayna-Yang](https://ETC5523-2025.github.io/assignment-4-packages-and-shiny-apps-Shayna-Yang)

Blog Post for BHAI analysis:

**Blog site: **[Germany’s Hidden Hospital Infection Burden: What the BHAI R Package Revealed](https://etc5523-2025.github.io/assignment-3-creating-a-blog-Shayna-Yang/posts/HAIs%20Burden%20in%20Germany/)

## License

This package is released under the MIT License. 
See the included LICENSE file for details.

## Author

Tzu-Hsuan, Yang (Author & Maintainer)

## Acknowledgements

- Package infrastructure: `usethis`, `devtools`, `roxygen2`

- Data wrangling: `dplyr`, `tidyr`, `scales`

- Visualisation: `ggplot2`, `plotly`, `shiny`

- Concept inspired by the BHAI methodology [(Zacher et al., 2019)](https://doi.org/10.2807/1560-7917.ES.2019.24.46.1900135)
